---
title: "Parents' Analysis"
author: "Wanting Tao"
date: "July 5, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
filename <- "C:/Users/TaoWa/Desktop/by_client_with_sequences.csv"
df1<- read.csv("C:/Users/TaoWa/Desktop/by_client_with_sequences.csv")

by_client <- read.csv("C:/Users/TaoWa/Desktop/by_client_with_sequences.csv")

library(ggplot2)
library(tidyverse)
library(DT)
library(forcats)
library(scales)

df1$collcode_y<-as.character(paste(df1$collcode,df1$year,sep="_"))
df1$y_collcode<-as.character(paste(df1$year,df1$collcode,sep="_"))

dff<-
  df1%>%
  rename(parents_college=q80)%>%
  mutate(parents_college=fct_recode(parents_college, 
                                    "yes" = "1", 
                                    "no" = "2", 
                                    "missing" = ".a", 
                                    "multiple" = ".b" ))%>%
  rename(gender=q72)%>%
  mutate(gender=fct_recode(gender, 
                                    "female" = "1", 
                                    "male" = "2", 
                                    "uncertain" = "3",
                                    "missing" = ".a", 
                                    "multiple" = ".b" ))
```

## parents' education * year

Using t_college to demonstrate how many apprentices from each college responsed the survey under different years.
t_parent shows how many apprentices' parents attended college/university before.
percentage of parents' education didn't change a lot.

```{r frequency tables}

table(dff$parents_college) # overall distribution for q80
ggplot(dff,aes(parents_college))+geom_bar()# total distubution for q80

t_q80 <- 
  table(dff$parents_college, df1$year) %>%
  as.data.frame() %>%
  rename(parents_college = Var1, year = Var2, surveys = Freq)

s_q80 <- 
  t_q80 %>%
  spread(key = year, value = surveys) %>%
  mutate(total = `13/14` + `14/15` + `15/16`) %>%
  arrange(desc(total))

s_q80%>%datatable()

ggplot(dff,aes(year,fill=parents_college)) +
  geom_bar(position="fill") + # plot for year, parents' education
  labs(title= "Graph 1", subtitle="response distribution by year", caption="no obvious changes during 3 years", y="response percentage")
```

## parents' college * college

The response "1" indicate parents attended the college/University education before, and "2" indicate not.
From the result table we can see only SENECA has more "2" than "1", which means most seneca students' parents didn't attend the education from college/University.

ALGO has most missing values.
SLAW has most proportion of parents_college = yes.
SENE has most proportion of parents_college = no.

```{r c_parent}
c_parent<-
  table(dff$collcode,dff$parents_college) %>%
  as.data.frame() %>%
  rename(collcode =Var1, parents_college = Var2)

s_parent<-
  c_parent %>%
  spread(key=parents_college,value=Freq)

s_parent%>%datatable()

p1<-
  ggplot(dff,aes(x=reorder(collcode,parents_college=="no",.asc=Ture),fill=parents_college)) +
  geom_bar(position="fill") + # position adjustments plot for parents' education in each college
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(title= "Graph 2", subtitle="college's parents rate", caption="order in parents didn't go to college/universities",x="college name",     y="response percentage")

print(p1)

```

## parents' education * gender
using g_parent to demonstrate gender* parents' education.
Female has higest proportion of parents_college = no.

```{r g_parent}
g_parent<-table(dff$gender,dff$parents_college) #q72*q80
prop.table(g_parent,1)   # row percentage

g_parent<-
  table(dff$gender,dff$parents_college) %>%
  as.data.frame() %>%
  rename(gender = Var1, parents_college= Var2)

sg_parent<-
  g_parent %>%
  spread(key=parents_college,value=Freq)
sg_parent%>% datatable()



ggplot(dff,aes(x=reorder(gender,parents_college=="no",.asc=Ture),fill=parents_college)) +
  geom_bar(position="fill") + # plot for q80 with gender
  labs(title= "Graph 3", subtitle="parents' education distribution for each gender", caption="female have highest satisfied rate",x="gender")
```

## top 10 trade by year

plot for college code, and overall questions
```{r top 10}
t_trade <- 
  table(dff$trade, dff$year) %>%
  as.data.frame() %>%
  rename(trade = Var1, year = Var2, surveys = Freq) #dataframe table by trade
  

s_trade <- 
  t_trade %>%
  spread(key = year, value = surveys) %>%
  mutate(total = `13/14` + `14/15` + `15/16`) %>%
  arrange(desc(total))

slice(s_trade,1:10)%>% datatable() #cut the top 10 trades in total

#subset observations
top<-c("309A","310S","306A","403A","310T","620C","433A","313A","308A","415A")
top_data<-filter(dff,trade%in%top)



ggplot(top_data,aes(x=reorder(trade,parents_college=="no",.desc=Ture),fill=parents_college)) +
  geom_bar(position="fill") + #distribution of parents' education for each trade
  labs(title= "Graph 4", subtitle="parents' rate distribution for each trade", caption="620C have highest proportion rate of parents no",x="trade")

```

## colleges response number by year

SENE only has year 13/14.
LACI only has year 13/14, 14/15
NIAG has most response number in 15/16.
```{r yc_parents,fig.width=10,fig.height=5}

yc_parents<-
  table(dff$y_collcode,dff$parents_college)%>%
  as.data.frame()%>%
  rename(y_collcode=Var1,parents_college=Var2)

yc_parents%>%datatable()


# college response nubmer change by year
ggplot(dff,aes(x=reorder(collcode,year=="15/16",.desc=Ture),fill=year)) +
  geom_bar(position="fill") + 
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(title= "Graph 5", subtitle="colleges response number during 3 years", caption="SENE only have one year record, LACI only have two years' records",x="college code", y="proportion by year")
```
## parents'education plus college plus year
## heat map: parents attended college/university by college and year

CAMB increase most: 55% to 71%, increase 16%.
ALGO decrease most: 70% to 38%, decrease 32%.
Highest rate appears in SSFL on year 14/15 for 77%.
lowest rate appears in ALGO on year 15/16 for 38%.

```{r heat map,fig.width=5,fig.height=10}
g_dff <-
  dff %>%
  group_by(collcode, year) %>%
  mutate(parents_yes = (parents_college=="yes" )) %>%
  summarise(n = n(),
            num_yes = sum(parents_yes),
            pct_yes = sum(parents_yes) / n())%>%
  mutate(p_yes= paste(round(100*pct_yes,2),"%",sep=""))

p11<-
  ggplot(g_dff,aes(year,collcode)) +
  geom_tile(aes(fill=pct_yes)) +
  geom_text(aes(label = percent(round(pct_yes, 2)), size = 2)) +
  scale_fill_gradient(low = "white", high = "coral") + 
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(title= "Graph 6", caption="SENE only have year 13/14, LACI doesn't have year 15/16", y ="college code")


print(p11)
```

## heat map: parents didn't attend college/university by college and year

LOYT increase most: 28% to 43%, increase 15%.
CAMB decrease most: 43% to 29%, decrease 14%.
Highest rate appears in SENE on year 13/14 for 50%.
lowest rate appears in ALGO on year 15/16 for 21%.

```{r heat map2,fig.width=5,fig.height=10}
g_dff <-
  dff %>%
  group_by(collcode, year) %>%
  mutate(parents_no = (parents_college=="no" )) %>%
  summarise(n = n(),
            num_no = sum(parents_no),
            pct_no = sum(parents_no) / n())%>%
  mutate(p_no= paste(round(100*pct_no,2),"%",sep=""))

p12 <-
  ggplot(g_dff, aes(year, collcode)) + 
  geom_tile(aes(fill = pct_no)) + 
  geom_text(aes(label = percent(round(pct_no, 2)), size = 2)) + #add label of each cell
  scale_fill_gradient(low = "white",  high = "coral") +                 #change color
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +            #change lable direction
  ggtitle("Graph 7") +            #add title
  labs(caption="SENE only have year 13/14, LACI doesn't have year 15/16",y="college code")                           #add caption, but can also add title, x-axis,y-axis
  
  
print(p12)
```