---
title: "Parents' education analysis"
author: "Wanting Tao"
date: "June 22, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
filename <- "C:/Users/TaoWa/Desktop/by_client_with_sequences.csv"
df1<- read.csv("C:/Users/TaoWa/Desktop/by_client_with_sequences.csv")

by_client <- read.csv("C:/Users/TaoWa/Desktop/by_client_with_sequences.csv")

library(ggplot2)
library(tidyverse)
library(DT)
library(forcats)

df1$collcode_y<-as.character(paste(df1$collcode,df1$year,sep="_"))
df1$y_collcode<-as.character(paste(df1$year,df1$collcode,sep="_"))

dff<-
  df1%>%
  rename(parents_college=q80)%>%
  mutate(parents_college=fct_recode(parents_college, 
                                    "yes" = "1", 
                                    "no" = "2", 
                                    "missing" = ".a", 
                                    "multiple" = ".b" ))%>%
  rename(gender=q72)%>%
  mutate(gender=fct_recode(gender, 
                                    "female" = "1", 
                                    "male" = "2", 
                                    "uncertain" = "3",
                                    "missing" = ".a", 
                                    "multiple" = ".b" ))
```


## parents' education * year

Using t_college to demonstrate how many apprentices from each college responsed the survey under different years.
t_parent shows how many apprentices' parents attended college/university before.
percentage of parents' education didn't change a lot.
```{r frequency tables}

table(dff$parents_college) # overall distribution for q80
ggplot(dff,aes(parents_college))+geom_bar()# total distubution for q80

t_q80 <- 
  table(dff$parents_college, df1$year) %>%
  as.data.frame() %>%
  rename(parents_college = Var1, year = Var2, surveys = Freq)

s_q80 <- 
  t_q80 %>%
  spread(key = year, value = surveys) %>%
  mutate(total = `13/14` + `14/15` + `15/16`) %>%
  arrange(desc(total))

s_q80%>%datatable()

ggplot(dff,aes(year,fill=parents_college))+geom_bar(position="fill")# plot for year, parents' education
```

## parents' college * college

The response "1" indicate parents attended the college/University education before, and "2" indicate not.
From the result table we can see only SENECA has more "2" than "1", which means most seneca students' parents didn't attend the education from college/University.

```{r c_parent}
c_parent<-
  table(dff$collcode,dff$parents_college) %>%
  as.data.frame() %>%
  rename(collcode =Var1, parents_college = Var2)

s_parent<-
  c_parent %>%
  spread(key=parents_college,value=Freq)

s_parent%>%datatable()

ggplot(dff,aes(parents_college,collcode))+geom_jitter()# plot for college, parents' education

ggplot(dff,aes(collcode,fill=parents_college))+geom_bar(position="fill")# position adjustments plot for parents' education in each college

ggplot(dff,aes(parents_college,fill=collcode))+geom_bar(position="fill")# position adjustments plot for each college' parents education.
```

## parents' education * gender
using g_parent to demonstrate gender* parents' education.For rows,"1" indicate female;"2" indicate male, "3" indicate other gender identify. Female aprrentices have similar proportion on parents' education, but male has an obvious larger proportion on higher parents' education.

```{r g_parent}
g_parent<-table(dff$gender,dff$parents_college) #q72*q80
prop.table(g_parent,1)   # row percentage

g_parent<-
  table(dff$gender,dff$parents_college) %>%
  as.data.frame() %>%
  rename(gender = Var1, parents_college= Var2)

sg_parent<-
  g_parent %>%
  spread(key=parents_college,value=Freq)
sg_parent%>% datatable()


ggplot(dff,aes(parents_college,gender))+geom_jitter() #plot for gender, parents' education
ggplot(dff,aes(parents_college,fill=gender))+geom_bar(position="fill")# plot for q80 with gender
```

## top 10 trade by year

plot for college code, and overall questions

```{r top 10}
t_trade <- 
  table(dff$trade, dff$year) %>%
  as.data.frame() %>%
  rename(trade = Var1, year = Var2, surveys = Freq) #dataframe table by trade
  

s_trade <- 
  t_trade %>%
  spread(key = year, value = surveys) %>%
  mutate(total = `13/14` + `14/15` + `15/16`) %>%
  arrange(desc(total))

slice(s_trade,1:10)%>% datatable() #cut the top 10 trades in total

#subset observations
top<-c("309A","310S","306A","403A","310T","620C","433A","313A","308A","415A")
top_data<-filter(dff,trade%in%top)

ggplot(top_data,aes(parents_college,trade))+geom_jitter() #plot for trade, parents' education
ggplot(top_data,aes(parents_college,fill=trade))+geom_bar(position="fill")# plot for q80 with trade
ggplot(top_data,aes(trade,fill=parents_college))+geom_bar(position="fill") #distribution of parents' education for each trade
```

## parents'education plus college plus year

Check parents' education for college by year
```{r yc_parents,fig.width=10,fig.height=5}

yc_parents<-
  table(dff$y_collcode,dff$parents_college)%>%
  as.data.frame()%>%
  rename(y_collcode=Var1,parents_college=Var2)

yc_parents%>%datatable()
# college response nubmer change by year
ggplot(dff,aes(collcode,fill=y_collcode))+geom_bar(position="fill")


# q80 distribution by college*year and change lable angle
yc_q80<-ggplot(dff,aes(collcode_y,fill=parents_college))+geom_bar(position="fill")
yc_q80 + theme(axis.text.x=element_text(angle=90,hjust=1))
```

