---
title: "survey_r"
author: "Wanting Tao"
date: "June 20, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
filename <- "C:/Users/TaoWa/Desktop/by_client_with_sequences.csv"
df1<- read.csv("C:/Users/TaoWa/Desktop/by_client_with_sequences.csv")

library(ggplot2)
library(tidyverse)
library(DT)
```


## responses, and parents * year

Using t_college to demonstrate how many apprentices from each college responsed the survey under different years.
t_parent shows how many apprentices' parents attended college/university before.
percentage of parents' education didn't change a lot.
```{r frequency tables}
t_college<-table(df1$collcode,df1$year)
t_college
table(df1$year) # response each year
t_parent<-table(df1$year,df1$q80)#q80 is parents college, parent's education * year
t_parent %>% datatable()
prop.table(t_parent,1) %>% datatable()  # row percentage
ggplot(df1,aes(q80,year))+geom_jitter()# plot for year, parents' education
```


## parents' college * college

The response "1" indicate parents attended the college/University education before, and "2" indicate not.
From the result table we can see only SENECA has more "2" than "1", which means most seneca students' parents didn't attend the education from college/University.

```{r c_parent}
c_parent<-table(df1$collcode,df1$q80) #q80 *college
c_parent
ggplot(df1,aes(q80,collcode))+geom_jitter()# plot for college, parents' education
```

## parents * gender
using g_parent to demonstrate gender* parents' education.For rows,"1" indicate female;"2" indicate male, "3" indicate other gender identify. Female aprrentices have similar proportion on parents' education, but male has an obvious larger proportion on higher parents' education.

```{r g_parent}
g_parent<-table(df1$q72,df1$q80) #q72*q80
g_parent
prop.table(g_parent,1)   # row percentage
ggplot(df1,aes(q80,q72))+geom_jitter() #plot for gender, parents' education

```

  
## top 10 trade by year

plot for college code, and overall questions

```{r top 10}
t_trade <- 
  table(df1$trade, df1$year) %>%
  as.data.frame() %>%
  rename(trade = Var1, year = Var2, surveys = Freq) 
  

s_trade <- 
  t_trade %>%
  spread(key = year, value = surveys) %>%
  mutate(total = `13/14` + `14/15` + `15/16`) %>%
  arrange(desc(total))

top_10<-slice(s_trade,1:10)
top_10

```

## Including Plots

plot for college code, and overall questions

```{r plots}

ggplot(df1,aes(q13s))+geom_bar()#plot for q13s

ggplot(df1,aes(q24s))+geom_bar()#plot for q24s

ggplot(df1,aes(q39s))+geom_bar()#plot for q39s

ggplot(df1,aes(q49s))+geom_bar()#plot for q49s
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
